---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r global_option}
library(knitr)
opts_chunk$set(tidy=FALSE, echo=TRUE, result="asis")
```

## Loading and preprocessing the data
```{r reading}
library(tidyverse)
act <- read_csv("activity.zip")
```
## What is mean total number of steps taken per day?
```{r histSteps}
daySteps <- act %>% group_by(date) %>% summarize(steps=sum(steps, na.rm=T ))
hist(daySteps$steps, breaks=12, main="Daily Steps", xlab="Steps")
```  

Mean and median of daily steps are

```{r mean_median_days}
sprintf("%6.0f", mean(daySteps$steps))  
sprintf("%6.0f", median(daySteps$steps))
```
respectively.

## What is the average daily activity pattern?
```{r timeSteps}
timeSteps <- act %>% group_by(interval) %>% summarize(aveSteps=mean(steps, na.rm=T ))
plot(timeSteps, type="l", xlab="time", ylab="average steps")
```

Maximum average of a day is in 
```{r maxAverage}
sprintf("%4.0f", timeSteps[which.max(timeSteps$aveSteps),1])
```

## Imputing missing values

There are `r sum(is.na(act$steps))` missing values in steps.

```{r missingValues}
sum(is.na(act$steps))
mean(is.na(act$steps))
```

I chose a very simple imputing methods of abandoning NA's.
The reason described below:

In this data, all NA's are all throughout the day. In other word a bunch of NA's are same date and there are no other data of the date.

```{r searchNA}
act %>% filter(is.na(steps)) %>% nrow
act %>% filter(is.na(steps)) %>% select(date) %>% unique()
```

there are 8 days which sums up to 2304 NA's and 

```{r}
288*8
```

As number of intervals of a day is (60/5) * 24 = `r (60/5) * 24`, they are all same date.

They are converted to 0 through
 
```{r daySteps}
daySteps <- act %>% group_by(date) %>% summarize(steps=sum(steps, na.rm=T ))
daySteps[1:10,]
```

Because all NA day contains no information and if we make NA as 0 it just increases number of zeroes.

As number of NA days are not very much

```{r numberofNA}
daySteps %>% summarize(mean(steps==0) ) 
```
I thought just excluding those NA days is a better idea.

=========================================================
```{r imputing}
acti <- act %>% filter(!is.na(steps))  # just abandon NA
dayStepsi <- acti %>% group_by(date) %>% summarize(steps=sum(steps))
hist(dayStepsi$steps, breaks=12, main="Imputed Daily Steps", xlab="Steps")

```

Mean and median of daily steps after imputing are

```{r mean_median_days_impute}
sprintf("%6.0f", mean(dayStepsi$steps))  
sprintf("%6.0f", median(dayStepsi$steps))
```
respectively.

Mean and median shifts to higher significantly after imputing.   
The effect of imputing is large therefore rationale is indispensable.

## Are there differences in activity patterns between weekdays and weekends?

```{r steps_week}
Sys.setlocale("LC_TIME", "en_US.UTF-8")
wkd <- function(dt) {
    ifelse(weekdays(dt) %in% c("Saturday","Sunday"), "weekend", "weekday")
}
actW <- acti %>% mutate(workday=wkd(date))
timeStepsW  <- actW %>%
    filter(workday == "weekday") %>%
    group_by(interval) %>%
    summarize(aveSteps=mean(steps, na.rm=T ))
timeStepsWe  <- actW %>%
    filter(workday == "weekend") %>%
    group_by(interval) %>%
    summarize(aveSteps=mean(steps, na.rm=T ))

dat <- c(timeStepsWe, timeStepsW)
ttl <- c("weekend", "weekday")
par(mfrow=c(2, 1), mar=c(0,0,0,0), oma=c(4,4,0.5,0.5))    
par(tcl=-0.25)
for (i in 1:2) {
    if (i==1) { 
        plot(timeStepsWe, type="l", axes=FALSE, ylim=c(0,220))
    } else {
        plot(timeStepsW,  type="l", axes=FALSE, ylim=c(0,220))
    }       
    mtext(ttl[i], side = 3, line = -1, adj = 0.1, cex = 1.5, col = "grey40")
    axis(2, col = "grey40", col.axis="grey20", gap.axis=0.5 ,at = seq(0,220,100))
    if (i==2) { 
        axis(1, col = "grey40", col.axis="grey20" ,cex=0.7, at = seq(0,2500,600))
    }       
    box(col = "grey40")
}
mtext("Interval", side = 1, outer = TRUE, cex = 1.0, line = 2.2, col = "grey20")   
mtext("Number of steps", side = 2, outer = TRUE, cex = 1.0, line = 2.2, col = "grey20")

dev.copy(png, filename="figure/PA1_template.png", width=480, height=480)
dev.off()
```
